Dish: Distribution-oblivious Shell Scripting

### 2020-02-29 pldi/draft.tex
Finish a pass on 4 only leaving some notes for additions

\kk{I have to mention that in order for the transformations to happen,
  we transform all commands that read some of their inputs in sequence
  to commands that read from one input but are preceded by a cat.}

\kk{Move this assumption in the transformation: Nodes may read from
  more files than just the files of their input stream (resp. write to
  more files than their output stream).  However, these files are
  static (\eg used for configuration or logging) and do not represent
  streams of data, and therefore are not considered part of the
  dataflow graph.  For now, we assume that these static files can only
  be accessed by one node, thus ensuring that they do not interfere
  with the distributed implementation.
%% Because of the assumption in section Front End \ref{}, that each file
%% exists only once in the dataflow graph, we can safely assume that
%% reading and writing to these static files does not interfere with the
%% distributed implementation, and doesn't alter the behaviour of the
%% program.
}

Having a dataflow graph representation allows us to define
parallelization optimizations as graph transformations. Before
describing the different types of optimizations, we formalize the
intuition about stateless and pure commands that was described
earlier~\sx{cmd}. We focus on parallelizing commands that produce one
output and consume their inputs one at a time, with the possible
exception of static input files that are used for configuration. More
formally, we parallelize commands that can be represented with
functions of type $f : D\kstar \times [D\kstar] \to D\kstar$, where
the first argument represents the concatenation of the inputs that are
consumed one at a time and the second argument represents the static
inputs. Most stateless and pure commands fall in this category
\kk{Maybe have an argument here that we show in the evaluation that
  handling these commands is adequate for a big variety of data
  processing scripts, from several microbenchmarks to some
  large-realistic macrobenchmarks}.
%% \nv{TODOs
%%   (i) highlight earlier the antithesis between the formalisms in this section and the informalities of the previous (me)
%% }

@@

\paragraph{Parallelization Transformations}
%
Based on these equations, we can define a maximal \emph{node split
  operation} on a stateless node $v$ with $n$ input ordered edges and
an output edge to node $v'$, to replace $v$ with $n$ new nodes, giving
each one of them one of the incoming edges of $v$, and routing them in
order to point to $v'$. Since each incoming edge represents a stream
of data $x_i : D^*$, and the only behaviour of a dataflow graph is its
output, this optimization $ v(x_1.x_2...x_n) \Rightarrow
v(x_1).v(x_2)...v(x_n)$ can be shown to preserve the behaviour of the
graph. Furthermore, it can be straightforwardly extended to pure
commands that can be implemented by a map-reduce pair $(m, r)$ as $
v(x_1.x_2...x_n) \Rightarrow r(m(x_1).m(x_2)...m(x_n))$.

@@

\kk{We should move FIgure 4 here}

\kk{I think that we should probably add the ``new'' split optimization
  (and maybe any others that we develop until then) here.}

@@

Given a dataflow graph representing a shell program, the \sys optimizer outputs an optimized dataflow graph in which data parallelism opportunities have been exploited.
More concretely, the
optimizer starts by splitting input files in several chunks, given the
desirable output dataflow graph width. It then starts from the source
nodes of the graph, iteratively performing the graph transformations
that were described in \Cref{ir:transformations}. If the incoming
edges of the node are less than the desirable graph width, it performs
edge splits, and then it performs maximal node splits. The final graph
can then be implemented by spawning a process for each node, and
redirecting the outputs according to the graph edges. An example
transformation can be seen in \Cref{fig:optimizer-transformation}.

\begin{figure}[t]
\centering
\includegraphics[width=0.40\columnwidth]{\detokenize{./figs/dish_g1.pdf}}
\caption{
  \textbf{Example graph transformation.}
  The optimizer splits a stateless node to utilize
  available data parallelism.
}
\vspace{-10pt}
\label{fig:optimizer-transformation}
\end{figure}

>>>>

%% \kk{Move this assumption in the transformation: Nodes may read from
%%   more files than just the files of their input stream (resp. write to
%%   more files than their output stream).  However, these files are
%%   static (\eg used for configuration or logging) and do not represent
%%   streams of data, and therefore are not considered part of the
%%   dataflow graph.  For now, we assume that these static files can only
%%   be accessed by one node, thus ensuring that they do not interfere
%%   with the distributed implementation.
%% %% Because of the assumption in section Front End \ref{}, that each file
%% %% exists only once in the dataflow graph, we can safely assume that
%% %% reading and writing to these static files does not interfere with the
%% %% distributed implementation, and doesn't alter the behaviour of the
%% %% program.
%% }

Having a dataflow graph representation allows us to define
parallelization optimizations as graph transformations. Before
describing the different types of optimizations, we formalize the
intuition about stateless and pure commands that was described
earlier~\sx{cmd}. We focus on parallelizing commands that produce one
output and consume their inputs one at a time, with the possible
exception of static input files that are used for configuration. More
formally, we parallelize commands that can be represented with
functions of type $f : D\kstar \times [D\kstar] \to D\kstar$, where
the first argument represents the concatenation of the inputs that are
consumed one at a time and the second argument represents the static
inputs. When translating a command that consumes its input streams one
at a time, a \ttt{cat} node is introduced before it to concatenate the
streams and its output is transferred to the command input. \kk{Maybe
  have an argument here that we show in the evaluation that handling
  these commands is adequate for a big variety of data processing
  scripts (and most stateless and pure commands), from several
  microbenchmarks to some large-realistic macrobenchmarks.}.
%% \nv{TODOs
%%   (i) highlight earlier the antithesis between the formalisms in this section and the informalities of the previous (me)
%% }

@@

\paragraph{Parallelization Transformations}
%
Based on these equations, we can define a \emph{node parallelization
  transformation} on a stateless node $v$ that is preceded by a
\ttt{cat} with $n$ input streams and is followed by a node $v'$. The
transformation replaces $v$ with $n$ new nodes, routing each of the
$n$ input streams to one of them, and commutes the \ttt{cat} node
after them to concatenate their outputs and transfer them to
$v'$. Since each incoming edge represents a stream of data $x_i :
D^*$, and the only behaviour of a dataflow graph is its output, this
optimization $ v(x_1.x_2...x_n, s) \Rightarrow v(x_1, s).v(x_2,
s)...v(x_n, s)$ can be shown to preserve the behaviour of the
graph. Furthermore, it can be straightforwardly extended to pure
commands that can be implemented by a map-reduce pair $(m, r)$ as $
v(x_1.x_2...x_n, s) \Rightarrow r(m(x_1, s).m(x_2, s)...m(x_n, s),
s)$. A visual representation of this transformation can be seen in
\Cref{fig:parallelization-transformation}.

\begin{figure}[t]
\centering
\includegraphics[width=0.60\columnwidth]{\detokenize{./figs/dish_g1.pdf}}
\caption{
  \textbf{Stateless parallelization transformation.}
  The \texttt{cat} node is commuted with the stateless node
  to utilize available data parallelism.
  \TODO{Correct this diagram to show the cat node before
    the node and then the cat node after the stateless
    command after the optimization. Cat node should be
    different than a circle. Maybe like an ``and'' node
    in binary circuits?}
}
\vspace{-10pt}
\label{fig:parallelization-transformation}
\end{figure}

@@

\kk{I think that we should probably rewrite this paragraph with the
  ``new'' split optimization (and maybe any others that we develop
  until then) here.}

@@

Given a dataflow graph representing a shell program, the \sys optimizer outputs an optimized dataflow graph in which data parallelism opportunities have been exploited.
More concretely, the
optimizer starts by splitting input files in several chunks, given the
desirable output dataflow graph width. It then starts from the source
nodes of the graph, iteratively performing the graph transformations
that were described in \Cref{ir:transformations}. If the incoming
edges of the node are less than the desirable graph width, it performs
edge splits, and then it performs maximal node splits. The final graph
can then be implemented by spawning a process for each node, and
redirecting the outputs according to the graph edges.
